---
title: "Social_Media_Analysis_Sravya_Kanagala"
output: html_document
date: "2024-03-28"
---

```{r}
# Load required library
library(stats)
library(readr)
library(stats)

```

```{r}
social_media_cleaned_up<- read.csv("C:/Users/Sravya/OneDrive/Desktop/RUTGERS UNI WORK (ALL)/MULLTIVARIATE ANALYSIS/social_media_cleaned.csv")

social_media_number <- social_media_cleaned_up[, -1]
social_media_scaled <- scale(social_media_number)
```

#### **Questions aimed to be answered with the analysis**

###### *Perform PCA (20 points), Cluster Analysis (20 Points), and Factor Analysis (20 points) on your data*

###### *Summarize what each of these models is telling you about yourself and the class (30 points)*

###### *Provide a takeaway from the analysis (10 points)*


### **Performing different methods**
#### **Principal Component Analysis**

```{r cars}
pca_result <- prcomp(social_media_scaled, scale. = TRUE)
summary(pca_result)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
